# Reproducible Research: Peer Assessment 1


Loading library for xyplot
```{r}
library(lattice)
```

## Loading and preprocessing the data

Loading and preprocessing the data
```{r echo=TRUE}
activity<-read.csv("activity.csv")
```

Preprocessing the data
```{r echo=TRUE}
totalByDate<-aggregate(steps~date, data=activity, sum, na.rm=TRUE)
```



## What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day
```{r echo=TRUE}
hist(	totalByDate$steps, 
      col = "red",
      main = "Histogram of the total number of steps taken each day",
      xlab = "Number of steps taken per day",
      ylab = "Frequency" )
```

Mean total number of steps taken per day
```{r echo=TRUE}
mean(totalByDate$steps)
```

Median total number of steps taken per day
```{r echo=TRUE}
median(totalByDate$steps)
```

## What is the average daily activity pattern?

Preprocessing the data
```{r echo=TRUE}
meanByInterval<-aggregate(steps~interval, data=activity, mean, na.rm=TRUE)
```

Time series plot of the average number of steps taken, averaged across all days (y-axis) vs
the 5-minute interval (x-axis)
```{r echo=TRUE}
plot(	steps~interval, 
      data = meanByInterval,
      type = "l", 
      main = "Average daily activity pattern",
      xlab = "Interval", 
      ylab = "Mean")
```

Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?
```{r echo=TRUE}
maxsteps = max(activity$steps, na.rm=TRUE)
#activity[which.max(activity$steps),]$interval
meanByInterval[which.max(meanByInterval$steps), ]$interval
```

## Imputing missing values

Total number of missing values in the dataset
```{r echo=TRUE}
sum(is.na(activity$steps))
```

strategy for filling in all of the missing values in the dataset
```{r echo=TRUE}
GetSimulatedData<-function(interval) {
    meanByInterval[meanByInterval$interval==interval,]$steps
}
```

Create a new dataset that is equal to the original dataset but with the
missing data filled in.
```{r echo=TRUE}
activitySimulated<-activity
for(i in 1:nrow(activitySimulated)){
    if(is.na(activitySimulated[i,]$steps)){
        activitySimulated[i,]$steps<-GetSimulatedData(activitySimulated[i,]$interval)
    }
}
```

Histogram of the total number of steps taken each day
```{r echo=TRUE}
totalByDateSimulated<-aggregate(steps~date, data=activitySimulated, sum, na.rm=TRUE)
hist(	totalByDate$steps,
      col = "red",
      main = "Histogram of the total number of steps taken each day",
      xlab = "Number of steps taken per day",
      ylab = "Frequency" )
```

Mean total number of steps taken per day
```{r echo=TRUE}
mean(totalByDateSimulated$steps)
```

Median total number of steps taken per day
```{r echo=TRUE}
median(totalByDateSimulated$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels weekday and weekend indicating whether a given date is a weekday or weekend
day.
```{r echo=TRUE}
activitySimulated$wday <- ifelse( (as.POSIXlt(as.Date(activitySimulated$date))$wday-1 %% 7) >= 5,
                                  "weekend",
                                  "weekday" )							  
```

Panel plot containing a time series plot of the average number of steps taken, averaged
across all weekday days or weekend days (y-axis) vs 5-minute interval (x-axis)
```{r echo=TRUE}
totalByWeekdaySimulated<-aggregate(steps~interval+wday, data=activitySimulated, mean, na.rm=TRUE)
xyplot( steps~interval|factor(wday),
        data=totalByWeekdaySimulated,
        ylab="Number of steps",
        type="l",        
        aspect=1/2
        )
```